Thermodynamic assessment of the Kalina Split cycle

### 2013-08-08 paper_SPLIT#3.tex
Basic addition

\documentclass[final,times,5p]{elsarticle}

@@

%\usepackage{fleqn}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{nomencl}
\usepackage{framed} % Framing content
\usepackage{multicol} % Multiple columns environment
\usepackage{float}
\usepackage{rotating}
\usepackage{float}
\usepackage[pdftex,bookmarks=true]{hyperref}
\hypersetup{ 
    colorlinks,% 
    citecolor=black,% 
    filecolor=black,% 
    linkcolor=black,% 
    urlcolor=black 
}

\biboptions{square,sort&compress}

@@

\title{Thermodynamic assessment of the Kalina Split Cycle}

@@

The Kalina Split Cycle is a thermodynamic process for converting thermal energy into electrical power, (i) using the ammonia-water mixture as working fluid, like a conventional Kalina Cycle, and (ii) with a varying ammonia concentration during the heat extraction. This second feature results in an improved match between the heat source and working fluid temperature profiles, decreasing the entropy generation in the heat recovery system.  
The present work compares the thermodynamic performance of this power cycle with other configurations of the Kalina process, and investigates the impact of varying operating conditions. The results indicate that the Kalina Split Cycle with reheat presents an exergetic efficiency higher by 2.8\% points than a reference Kalina cycle with reheat, and by 4.3\% points without. The cycle performance is highly sensitive to its boundary conditions, varying by 9\% for a temperature variation of 30$^{\circ}$C for the cold reservoir and 100$^{\circ}$C for the stack. This analysis also pinpoints the large irreversibilities in the low-pressure turbine and condenser, which have an efficiency defect of 9.5-13\% and 7.2-13.7\%, respectively, and indicates a reduction of the exergy destruction by about 23\% in the heat recovery system.

@@

\begin{keyword}
Kalina Split Cycle \sep Energy analysis \sep Exergy assessment \sep Waste Heat Recovery 
\end{keyword}

@@

Waste heat recovery (WHR) systems are able to generate mechanical power and electricity without any fuel input and associated CO$_2$ emissions. Hence, with rising fuel prices and increased environmental awareness, motivation is growing for integrating these systems to improve the energy efficiency of various processes. An example is a WHR system for large marine engines in the present study. Organic Rankine cycles (ORC) and Kalina cycles are among the most frequently proposed alternatives to steam Rankine cycle WHR systems. At the scale of application studied in the present work, i.e. a net power output of 1-5 MW, both power cycles may be viable \cite{Tchanche20113963,Jonsson2001c}. Bombarda et al. \cite{Bombarda2010b} compared the two processes for a heat source temperature level of 346$^\circ$C and found that both cycles, when optimised, produced almost equal net power outputs. While the present study does not directly compare the ORC with the Kalina cycle, this work is based on the boundary conditions used in the work of Bombarda et al. \cite{Bombarda2010b}, in order to make comparisons about the power cycle performance for the same application.

Energy can neither be created nor destroyed, and an energy analysis illustrates the energy transformations and flows throughout the system under study. On the opposite, exergy is not conserved in real processes, illustrating therefore the locations, causes and magnitudes of the thermodynamic irreversibilities taking place \cite{BejanAdrian;TsatsaronisGeorge;Moran1996,Bejan2006}. Exergy destruction also accounts for the additional exergetic fuel required because of the system imperfections \cite{Wall1988,Kotas1980,Kotas1980a,Kotas1995}. A few studies on the thermodynamic performance of the Kalina cycle exist.

The present paper presents and evaluates a novel power generation cycle, called the Kalina Split Cycle. This process is also based on the ammonia-water mixture as working fluid, like the conventional Kalina Cycle, but is characterised by a varying ammonia concentration in the heat recovery system. This results in a smaller entropy generation in the heat transfer process, and potentially in a higher exergetic efficiency of the complete power cycle. This concept was briefly mentioned in the work of Kalina, and a system analysis was previously conducted by the same authors to identify the governing mechanisms of this process. The literature appears to contain little, if none, on the thermodynamic performance of such cycle, and this study aims at closing this gap, following these three objectives:

@@

	\item estimation of the cycle potential, in terms of energy and exergy efficiencies, compared to a conventional Kalina cycle, with and without reheat;
	\item analysis of the plant inefficiencies and of the exergy destruction trends;
	\item evaluation of the impact of the boundary conditions (heat source and cold reservoir temperatures) on the performance of this system;

@@

Section \ref{sec:system_description} presents the design of the Kalina Split Cycle system and the methods used in this work are reported in Section \ref{sec:methods}. The results are presented in Section \ref{sec:results} and are criticised further in Section \ref{sec:discussion}. Concluding remarks are outlined in Section \ref{sec:conclusion}.

@@

\section{System description}
\label{sec:system_description} 
The reference Kalina process layout and the process conditions used throughout were the same as those presented in the work of Bombarda et al. \cite{Bombarda2010b}. Both the reference cycle and the Split-cycle were evaluated with and without using reheat in the turbine, in order to determine the influence of this technique on the processes.

In the following, the solution concentration running through the turbine is referred to as the \emph{working solution}, and the terms \emph{lean} and \emph{rich} refer to a low and a high concentration of ammonia in the solution, respectively.

\subsection{Reference Kalina cycle}
\label{sec:kalina_cycle}
\begin{figure*}[htbp]
\centering
\includegraphics[width=0.9\linewidth]{Drawing_Kalina_Baseline.pdf}
\caption{Sketch of the Kalina reference process with reheat}
\label{fig:kalina_cycle}
\end{figure*}

Figure \ref{fig:kalina_cycle} illustrates the flow diagram of the reference Kalina process with reheat. Starting from (21) to (1), the preheated working fluid is evaporated and superheated in the boiler before it enters the turbine (3). In the process layout that includes the reheat technique, the outlet stream from the turbine (3') is heated in the boiler before entering (3'') a second turbine. When reheat is not included in the process, stream (3) runs directly from the turbine outlet (4) to Recuperator 1. From the stream (4) heat is transferred to the stream (10) in Recuperator 1. The stream (5) is then mixed with an ammonia lean stream from the separator (15) to form a leaner solution. This solution is condensed (7) and after being pumped to an intermediate pressure level, the stream (8) is divided into two streams (9) and (17). The stream (9) is heated in Recuperator 2 and in Recuperator 1 to a partially evaporated state. It then enters the separator which separates the stream into a lean liquid (12) and a very rich vapour (13). Heat from stream (13) is used to preheat stream (20) in Recuperator 3, and the stream (16) is then mixed with a leaner solution (17) to form the working solution (18). This stream is finally condensed and pumped to the boiler pressure.

\subsection{Kalina Split-cycle}
\label{sec:split_cycle}
\begin{figure*}[htbp]
\centering
\includegraphics[width=1.0\linewidth]{Drawing_Kalina_Split.pdf}
\caption{Sketch of the Kalina Split-cycle process}
\label{fig:split_cycle}
\end{figure*}

Figure \ref{fig:split_cycle} illustrates the flow diagram of the Split-cycle process. To maintain focus on the special split stream boiler, the Split-cycle configuration modelled in this work was designed to have a minimum number of components needed for evaluating the concept. Hence, the SC presented here is based on the same principles as the reference Kalina cycle, with some important differences.

Two streams of different ammonia concentration enter the boiler, a rich stream (25) and a lean stream (31). Before being mixed (Mixer 4), the rich stream is fully evaporated, and the lean stream is heated to the bubble point state. The aim of this arrangement is to lower the temperatures in the overall process going from liquid (25,31) to vapour (2). To be able to produce these two streams with the desired concentrations and mass flow rates, an additional mixing subsystem, consisting of three splitters and two mixers, is also needed.

The gradient of the evaporation temperature curve can to some degree be adjusted to the temperature profile of the heat source by selecting the optimal composition of each of the two streams, as illustrated in Fig. \ref{fig:theory}. The line from (25,31) to the point ($T_{r,b}$)\nomenclature[S]{r}{rich ammonia concentration} \nomenclature[S]{b}{bubble point}represents the preheat stage. From here to points (1), (2) and (3) the fully drawn line represents the heat transfer when using the Split-cycle configuration. The upper dashed line represents how the heat transfer would be if the two streams were combined into a single stream. The point ($T_{b}$) signifies the bubble point of the combined stream, and it is clear that the temperature difference at the pinch point is much smaller and possibly violated in this case. The lower dashed line from ($T_{r,b}$) through (2a) to (2), shows how the heat transfer would occur if only the rich stream concentration was used. Evaporation would take place at lower temperatures possibly leading to a lower thermal efficiency.

@@

\caption{Sketch of T-Q diagram of the heat recovery system to explain the Split-cycle}
\label{fig:theory}
\end{figure*}

Kalina argued \cite{Kalina1986a} that the state of the rich stream at point (33) (Fig. \ref{fig:split_cycle}) should ideally be at the dew point and that the lean stream at point (32) should be at the bubble point, before the mixing of the two streams. The two streams should also have similar temperatures and pressures in order to minimise the entropy generation in this section of the boiler. We decided to adopt these constraints without further analysis to focus on the full process analysis. However, an entropy generation analysis is planned for future work to verify this claim. In the following, these conditions are referred to as the SC boiler constraints.

A direct consequence of the SC boiler constraints is that, once the ammonia concentration of one of the streams, (25) or (31), has been chosen, the concentration of the other is fixed in order to satisfy the equilibrium conditions. Additionally, when the boiler pressure and the concentration in one stream are chosen, the temperature of the working fluid streams out of evaporator 1 is fixed. Both are illustrated in the example shown in Fig. \ref{fig:bubbleDew}. As shown, the mixing temperature decreases and the lean stream concentration increases, as the rich stream concentration increases.

\begin{figure}[htpb]
\centering
\begin{tikzpicture}[scale=0.9]

\begin{axis}[xlabel={Rich stream ammonia concentration},
	%xmin=0.75, xmax=1.0,
	ymin=0.2,
	axis x line*=bottom,
	axis y line*=left,
	ylabel={Lean stream ammonia concentration},
	ylabel near ticks,
	legend style={legend pos=south west,font=\small, draw=none} ,
	]

@@

\caption{Equilibrium conditions for Evaporator 1 outlet}
\label{fig:bubbleDew}
\end{figure}

@@

	\subsection{Process modelling, simulation and validation}

@@

	where $\dot{Q}_{k}$ and $\dot{W}_{cv}$ the time rates of energy transfer by heat and work ($\dot{Q}\ge0$ indicates heat transfer to the system, $\dot{W}\ge0$ work done by the system). The subscripts $in$ and $out$ denote the inlet and outlet of the system, $k$ the boundary of the component of interest and $cv$ the control volume. $h$ is the specific enthalpy of the stream under consideration, and can be defined as the change in enthalpy observed when it is brought from its temperature and pressure to the reference conditions, and is decomposed into the reference elements in standard state.

@@

	The energy balance for the Kalina Split Cycle can be expressed as:

	\begin{equation}
	 \dot{Q}_{heat}=\dot{W}+\dot{Q}_{cool}
	\end{equation}

@@

	In the absence of nuclear, magnetic and electrical interactions, the exergy associated with a stream of matter is a function of its physical $e^{ph}$, chemical $e^{ch}$, kinetic $e^{kn}$ and potential $e^{pt}$ components \cite{BejanAdrian;TsatsaronisGeorge;Moran1996}. Therefore, on a unit mass basis, the specific exergy of a material stream is expressed as:

@@

	Physical exergy accounts for temperature and pressure differences from the environmental state and is defined as:

@@

		\item The exergetic efficiency $\varepsilon$ of a given component or sub-system $k$, which reflects its thermodynamic performance. It is defined as the ratio of the product exergy to the fuel exergy. The product exergy $\dot{E}_{p,k}$ represents the desired effect of a given thermodynamic transformation or process, while the fuel exergy $\dot{E}_{f,k}$ represents the resources expended in this component/sub-system to generate this desired result.

@@

	Fuel and product exergies are not always equal to the exergy flows entering $\dot{E}_{in,k}$ and leaving $\dot{E}_{out,k}$ the component/sub-system: they represent the exergy flows that are useful (product) or used (fuel). The definitions of exergetic fuels and products for the components existing in power cycle processes are introduced and discussed in Kotas \cite{Kotas1980,Kotas1980a,Kotas1995} and in Bejan et al. \cite{BejanAdrian;TsatsaronisGeorge;Moran1996}.

@@

	The exergy accounting for the Kalina Split Cycle can therefore be expressed as:

	\begin{equation}
		\dot{E}^{Q}_{heat}=\dot{E}^{W}+\dot{E}^{Q}_{cool}+\dot{E}_{d}
	\end{equation}

@@

\nomenclature[S]{cool}{cooling medium}

>>>>

\documentclass[final,times,3p]{elsarticle}

@@

%\usepackage{fleqn}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{nomencl}
\usepackage{framed} % Framing content
\usepackage{multicol} % Multiple columns environment
\usepackage{float}
\usepackage{rotating}
\usepackage{float}
\usepackage[pdftex,bookmarks=true]{hyperref}
\hypersetup{ 
    colorlinks,% 
    citecolor=black,% 
    filecolor=black,% 
    linkcolor=black,% 
    urlcolor=black 
} 
\usepackage{pgfplots, pgfplotstable}
\biboptions{square,sort&compress}

@@

\title{Thermodynamic assessment of the Kalina Split cycle}

@@

The Kalina Split cycle is a thermodynamic process for converting thermal energy into electrical power, (i) using the ammonia-water mixture as working fluid, like a conventional Kalina cycle, and (ii) with a varying ammonia concentration during the heat extraction. This second feature results in an improved match between the heat source and working fluid temperature profiles, decreasing the entropy generation in the heat recovery system.  
The present work compares the thermodynamic performance of this power cycle with other configurations of the Kalina process, and investigates the impact of varying operating conditions. \textcolor{green}{The results indicate that the Kalina Split cycle with reheat presents an exergetic efficiency higher by 2.8\% points than a reference Kalina cycle with reheat, and by 4.3\% points without. The cycle performance is highly sensitive to its boundary conditions, varying by 9\% for a temperature variation of 30$^{\circ}$C for the cold reservoir and 100$^{\circ}$C for the stack. This analysis also pinpoints the large irreversibilities in the low-pressure turbine and condenser, which have an efficiency defect of 9.5-13\% and 7.2-13.7\%, respectively, and indicates a reduction of the exergy destruction by about 23\% in the heat recovery system.}

@@

\begin{keyword}
Kalina Split cycle \sep Energy analysis \sep Exergy assessment \sep Waste Heat Recovery 
\end{keyword}

@@

\textcolor{red}{Waste heat recovery (WHR) systems are able to generate mechanical power and electricity without any fuel input and associated CO$_2$ emissions. Hence, with rising fuel prices and increased environmental awareness, motivation is growing for integrating these systems to improve the energy efficiency of various processes. An example is a WHR system for large marine engines in the present study. Organic Rankine cycles (ORC) and Kalina cycles are among the most frequently proposed alternatives to steam Rankine cycle WHR systems. At the scale of application studied in the present work, i.e. a net power output of 1-5 MW, both power cycles may be viable \cite{Tchanche20113963,Jonsson2001c}. Bombarda et al. \cite{Bombarda2010b} compared the two processes for a heat source temperature level of 346$^\circ$C and found that both cycles, when optimised, produced almost equal net power outputs. While the present study does not directly compare the ORC with the Kalina cycle, this work is based on the boundary conditions used in the work of Bombarda et al. \cite{Bombarda2010b}, in order to make comparisons about the power cycle performance for the same application.}

Energy can neither be created nor destroyed, and an energy analysis illustrates the energy transformations and flows throughout the system under study. On the opposite, exergy is not conserved in real processes, illustrating therefore the locations, causes and magnitudes of the thermodynamic irreversibilities taking place \cite{BejanAdrian;TsatsaronisGeorge;Moran1996,Bejan2006}. Exergy destruction also accounts for the additional exergetic fuel required because of the system imperfections \cite{Wall1988,Kotas1980,Kotas1980a,Kotas1995}. Several studies on the thermodynamic performance of the Kalina cycle exist.

The present paper presents and evaluates a novel power generation cycle, called the Kalina Split cycle. This process is also based on the ammonia-water mixture as working fluid, like the conventional Kalina cycle, but is characterised by a varying ammonia concentration in the heat recovery system. This can result in a smaller entropy generation in the heat transfer process, and potentially in a higher exergetic efficiency of the complete power cycle. \textcolor{blue}{Shall we emphasise somewhere that we verify the statement smaller entropy generation in the boiler involves higher exergetic efficiency of the cycle?} This concept was briefly mentioned in the work of Kalina, and a system analysis was previously conducted by the same authors to identify the governing mechanisms of this process. The literature appears to contain little, if none, on the thermodynamic performance of such cycle, and this study aims at closing this gap, following these three objectives:

@@

	\item estimation of the cycle potential, in terms of energy and exergy efficiencies, compared to a conventional Kalina cycle, with and without reheat;
	\item analysis of the plant inefficiencies and of the exergy destruction trends;
	\item evaluation of the effect of the boundary conditions (heat source and cold reservoir temperatures) on the system performance.

@@

Section \ref{sec:system_description} presents the design of the Kalina Split cycle system and the methods used in this work are reported in Section \ref{sec:methods}. The results are presented in Section \ref{sec:results} and are criticised further in Section \ref{sec:discussion}. Concluding remarks are outlined in Section \ref{sec:conclusion}.

@@

\section{System description}
\label{sec:system_description}

\subsection{Reference Kalina cycle}
\label{sec:kalina_cycle}

The Kalina cycle is similar in principle to the Rankine cycle, in which heat is supplied to a closed process loop and thermal energy is transformed into mechanical work. The main difference lies in the properties of the working fluid, which is an ammonia-water mixture in the Kalina cycle. This multicomponent mixture is zeotropic, which means that the vapor and liquid phases during evaporation do not present the same composition, At constant pressure, the boiling temperature changes during the heat transfer process, at the difference of pure substances, which have a constant evaporation temperature.

This temperature glide allows a better matching between the temperature profiles of the heat source and receiver, causing therefore a smaller exergy destruction during the heat transfer process. The range of the temperature glide can be adjusted by modifying the ammonia and water fractions of the working solution, and the Kalina cycle may therefore be suitable for application in both low- and medium-temperature heat recovery applications.

Several configurations of this thermodynamic cycle exist, and the layout considered in this work (Figure \ref{fig:kalina_cycle}) is the one presented in the study of Bombarda et al. \cite{Bombarda2010b} and was studied by Marston \cite{Marston1990,Marston1995} and El-Sayed \cite{El-Sayed1985}. The terms `rich' and `lean'  imply ammonia-rich and ammonia-lean in the rest of this work. Four different ammonia concentrations can be found in the conventional Kalina cycle: `very rich' (points 13 and 16), `rich' (points 1 to 5, and 18 to 21), `medium' (points 6 to 11, and 17) and  `lean' (points 12, 14 and 15).

\begin{figure*}[htbp]
\centering
\includegraphics[width=0.9\linewidth]{Drawing_Kalina_Baseline.pdf}
\caption{Flowchart of the Kalina cycle process}
\label{fig:kalina_cycle}
\end{figure*}

The ammonia-rich mixture (1) enters the heat recovery system where it is evaporated (2) and superheated (3) by heat recovery. The working solution is expanded to a low pressure level (4), generating shaft work, and enters a first heat exchanger (Recuperator 1) where it is cooled and partially condensed (5). Reheat may be implemented, implying that the stream exiting the turbine re-enters the boiler at an intermediate pressure (3') and is heated (3'') to the same temperature as (3), before being expanded in a second turbine. Stream (5) is mixed with the lean solution (15), resulting in a working solution (6) with a higher condensing temperature. It is condensed (7), pumped to an intermediate pressure level (8), and split into two streams (9) and (17). Stream (9) is heated in a succession of two heat exchangers (Recuperator 2 and Recuperator 1) and enters (11) the vapour-liquid separator to form a lean liquid (12) and a very rich vapour (13). The lean liquid is cooled (14) and throttled to the lowest pressure level (15) of this power cycle.  The very rich vapour is cooled (16) and diluted with (17) into (18) in a second mixer. This working solution is condensed (19) and pumped to the highest pressure level (20), i.e. the boiler pressure, heated in two steps (21), and (1), by recovering heat from the waste heat source.

\subsection{Kalina Split-cycle}
\label{sec:split_cycle}

The Kalina Split cycle is fundamentally similar to the Kalina cycle (Figure \ref{fig:split_cycle}), both using the ammonia-water mixture. The main difference lies in the change of ammonia concentration in the evaporation process, which involves a more complex splitting and mixing arrangement to achieve the desired ammonia concentrations. Five different concentrations can be found in the Kalina Split cycle and are denoted `very rich' (points 18, 19 and 26), `rich' (points 20 to 25, and 33) , `basic' (points 1 to 5), `lean' (points 27 to 32), and  `very lean' (points 11 to 16), in reference to their ammonia content.

\begin{figure*}[htbp]
\centering
\includegraphics[width=1.0\linewidth]{Drawing_Kalina_Split.pdf}
\caption{Flowchart of the Kalina Split cycle process}
\label{fig:split_cycle}
\end{figure*}

At the difference of the Kalina cycle, two streams of different ammonia concentration enter the boiler (Figure \ref{fig:theory}, left), an ammonia-rich (25) and an ammonia-lean (31). The first one is fully evaporated (33) while the second one stays in liquid form, but is heated to its bubble point temperature (32), They are mixed into (1): the resulting solution is an ammonia-water solution in vapour-liquid equilibrium, which is evaporated (2) and superheated (3).

The preheating stage is illustrated by the (25,31)--$T_{r,b}$ line: the lean and rich streams are sensibly heated to the bubble point temperature, recovering energy from the heat source. The dashed upper line $T_{r,b}$--$T_b$ illustrates the heat transfer-temperature path if the rich and lean streams enter the boiler as a single stream, as in the conventional Kalina cycle. The dashed lower line $T_{r,b}$--$T_{2a}$ represents the T -Q path if a single stream with the same ammonia concentration as the rich stream enters the boiler. \textcolor{blue}{What about the lean stream? I guess it would be under as it would not start yet to evaporate, but maybe we should write a bit on that too?} The full line $T_{r,b}$--$T_{1}$ corresponds to the Split Cycle configuration, with two streams of different concentrations.

The difference between the paths $T_{r,b}$--$T_b$ and $T_{r,b}$--$T_{1}$ suggests that the Kalina Split cycle offers more flexibility, as the concentrations of the two streams entering the boiler can be adjusted to prevent a too small temperature difference between the heat source and receiver. The difference between the paths $T_{r,b}$--$T_{1}$ and $T_{r,b}$--$T_{2a}$ indicates that the split cycle configuration results in a better match of the enthalpy-temperature curves, and therefore in a smaller entropy generation in the boiler. In all cases, the minimum temperature difference, i.e. the pinch point location, is found in the evaporation process, i.e after the preheating and before the mixing of the rich and lean streams.

\nomenclature[S]{r}{rich ammonia concentration} \nomenclature[S]{b}{bubble point}

This work assumes that the rich and lean streams are mixed at same temperature and pressure to reduce the entropy generation due to mixing effects. In addition, Kalina \cite{Kalina1986a} suggested that the rich stream should be at dew point conditions, while the lean stream should be at bubble point ones. The vapour and liquid phases of each stream are therefore at their equilibrium concentrations, which would minimise the entropy generation in the mixing process. However, such constraint fixes the mixing temperature, as the ammonia concentration of the lean and rich streams are interdependent. The actual mixing point has little influence on the overall performance of the system, since the pinch point is indeed located either at the bubble temperature of the ammonia-rich stream $T_{r,b}$--$T_{b}$ or at the outlet of the superheater $T_{3}$--$T_{34}$.

@@

\end{axis}
\end{tikzpicture}
\hfill
\begin{tikzpicture}[scale=0.9]

\begin{axis}[xlabel={Rich stream (25) NH$_3$-concentration},
	%xmin=0.75, xmax=1.0,
	ymin=0.2,
	axis x line*=bottom,
	axis y line*=left,
	ylabel={Lean stream (31) NH$_3$-concentration},
	ylabel near ticks,
	legend style={legend pos=south west,font=\small, draw=none} ,
	]

@@

\caption{Split Cycle Mechanisms: T-Q diagram of the heat recovery system (\emph{left}) and Temperature-NH$_3$ concentration dependency (\emph{right})}
\label{fig:theory}
\end{figure*}

@@

\subsection{Baseline case}

This study builds on a baseline case adapted from the work of Bombarda et al. \cite{Bombarda2010b}, where the integration of a bottoming Kalina cycle is studied, using exhaust gases from diesel engines as heat source. The Kalina cycle is compared to the Kalina Split cycle, considering one reheat stage as a possible option (Table \ref{tab:system_baseline}).

\begin{table}[htbp]
\scriptsize
  \centering
  \caption{System parameters (baseline case)}
    \begin{tabular}{llr}
    \toprule
    Parameter & Description  & Value \\
    \midrule
     $T_{34}$ & Heat source inlet temperature & 346 $^{\circ}$C \\
     $T_{37}$     & Heat source outlet temperature & 127.7 $^{\circ}$C\\
     $T_{cw}$    & Cooling water inlet temperature & 25 $^{\circ}$C\\
     $T_{34}-T_{3}$  & Superheater approach & 16 $^{\circ}$C\\
     $\Delta T_{evap}$     & Minimum temperature difference (evaporators) & 21.9 $^{\circ}$C \\
     $\Delta T_{rec}$    & Minimum temperature difference (recuperators) & 5 $^{\circ}$C \\
     $\Delta T_{cond}$   & Minimum temperature difference (condenser) & 4.5 $^{\circ}$C \\
     $\eta_{t,pol}$    & Turbine polytropic efficiency & 70.5$\%$\\
     $\eta_{t,mec}$    & Turbine mechanical efficiency & 96$\%$\\
     $\eta_{pp}$    & Pump efficiency & 70$\%$\\
     $\eta_{dr}$  & Driver efficiency & 95$\%$\\
    \bottomrule
    \end{tabular}%
  \label{tab:system_baseline}%
\end{table}

\pgfplotstableread[col sep=comma,header=false]{
Category 1,6,0,0
Category 2,2,2,0
Category 3,3.5,2.5,4
Category 4,3.5,2.5,4
Category 5,3.5,2.5,4
Category 6,3.5,2.5,4
Category 7,3.5,2.5,4
Category 8,3.5,2.5,4
Category 9,3.5,2.5,4
Category 10,3.5,2.5,4
Category 11,3.5,2.5,4
}\data

\pgfplotstablecreatecol[
    create col/expr={
        \thisrow{1} + \thisrow{2} + \thisrow{3} 
    }
]{sum}{\data}

\pgfplotsset{
    percentage plot/.style={
        point meta=explicit,
    every node near coord/.append style={
        align=center,
        text width=1cm
    },
        nodes near coords={
        \pgfmathtruncatemacro\iszero{\originalvalue==0}
        \ifnum\iszero=0
            \pgfmathprintnumber{\originalvalue}$\,\%$\\ \pgfmathprintnumber[fixed zerofill,precision=1]{\pgfplotspointmeta}
        \fi},
    nodes near coords align=vertical,
        yticklabel=\pgfmathprintnumber{\tick}\,$\%$,
        ymin=0,
        ymax=100,
        enlarge y limits={upper,value=0.18},
    visualization depends on={y \as \originalvalue}
    },
    percentage series/.style={
        table/y expr=\thisrow{#1}/\thisrow{sum}*100,table/meta=#1
    }
}

%%% TRY TO PLOT COMPARISON %%% 
\begin{tikzpicture}
\begin{axis}[
    axis on top,
    width=\textwidth,
    percentage plot,ybar,bar width=0.1cm,
    enlarge x limits=0.25,
    symbolic x coords={Category 1,Category 2,Category 3,Category 4,Category 5,Category 6,Category 7,Category 8,Category 9,Category 10,Category 11},
    xtick=data
]
\addplot table [percentage series=1] {\data};
\addplot table [percentage series=2] {\data};
\addplot table [percentage series=3] {\data};
%\addplot table [percentage series=4] {\data};
\end{axis}
\end{tikzpicture}

@@

	where $\dot{Q}_{k}$ and $\dot{W}_{cv}$ the time rates of energy transfer by heat and work. The convention adopted in this work is that $\dot{Q}\ge0$ indicates heat transfer to the system and $\dot{W}\ge0$ work done by the system. The subscripts $in$ and $out$ denote the inlet and outlet of the system, $k$ the boundary of the component of interest and $cv$ the control volume. $h$ is the specific enthalpy of the stream under consideration, and can be defined as the change in enthalpy observed when it is brought from its temperature and pressure to the reference conditions, and is decomposed into the reference elements in standard state.

@@

	The energy balance for the Kalina Split cycle can be expressed as:

	\begin{equation}
	 \dot{Q}_{heat}=\dot{W}+\dot{Q}_{cw}
	\end{equation}

@@

	The exergy associated with a stream of matter is a function of its physical $e^{ph}$, chemical $e^{ch}$, kinetic $e^{kn}$ and potential $e^{pt}$ components \cite{BejanAdrian;TsatsaronisGeorge;Moran1996} and is expressed, on a specific mass basis, as:

@@

	Physical exergy accounts for temperature and pressure differences from the environmental conditions and is defined as:

@@

		\item The exergetic efficiency $\varepsilon$ of a given component or sub-system $k$, which reflects its thermodynamic performance. It is defined as the ratio of the product exergy to the fuel exergy.

@@

	Fuel and product exergies are not always equal to the exergy flows entering $\dot{E}_{in,k}$ and leaving $\dot{E}_{out,k}$ the component/sub-system: the product exergy $\dot{E}_{p,k}$ represents the desired effect of a given thermodynamic transformation or process, while the fuel exergy $\dot{E}_{f,k}$ represents the resources expended in this component/sub-system to generate this desired result. The definitions of exergetic fuels and products for the components existing in power cycle processes are introduced and discussed in Kotas \cite{Kotas1980,Kotas1980a,Kotas1995} and in Bejan et al. \cite{BejanAdrian;TsatsaronisGeorge;Moran1996}.

@@

	The exergy accounting for the Kalina Split cycle can therefore be expressed as:

	\begin{equation}
		\dot{E}^{Q}_{heat}=\dot{E}^{W}+\dot{E}^{Q}_{cw}+\dot{E}_{d}
	\end{equation}

@@

\nomenclature[S]{cw}{cooling water}
